\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bbm, enumerate}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{note}[1][Note.]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\begin{document}

\title{Differential Geometry: Homework 1}
\author{Peter Kagey}

\maketitle

% -----------------------------------------------------
% First problem
% -----------------------------------------------------
\begin{problem}{1}
  Show that the induced topology
  (for a subset $X \subset Y$ of a topological space $Y$)
  and the quotient topology (for a surjection $X \twoheadrightarrow Y$ from a
  topological space $X$ onto a set $Y$)
  satisfy the axioms of a topological space.
\end{problem}

\begin{proof}
\end{proof}

% -----------------------------------------------------
% Second problem
% -----------------------------------------------------
\pagebreak

\begin{problem}{2}
  Show that the topological spaces $S^1 \subset \mathbb{R}^2$
  (with topology induced by the inclusion into $\mathbb{R}^2$)
  and $[0, 1]/\{0, 1\}$
  (with the quotient topology from the topology on $[0, 1] \subset \mathbb{R}$)
  are homeomorphic.
\end{problem}

\begin{proof}
\end{proof}

% -----------------------------------------------------
% Third problem
% -----------------------------------------------------
\pagebreak

\begin{problem}{3}
  Prove that $S^1$, with either topology considered above, is a topological
  manifold.
\end{problem}

\begin{proof}
\end{proof}

% -----------------------------------------------------
% Fourth problem
% -----------------------------------------------------
\pagebreak

\begin{problem}{4}
  Show that the derivative of a function
  $f: \mathbb{R}^n \rightarrow \mathbb{R}^m$
  if it exists at a point $a \in \mathbb{R}^m$ is unique.
\end{problem}

\begin{proof}
\end{proof}

% -----------------------------------------------------
% Fifth
% -----------------------------------------------------
\pagebreak

\begin{problem}{5}
  Produce, with proofs, examples of the following topological spaces which are
  not topological manifolds: \begin{enumerate}[a)]
    \item A space $X$ which is locally Euclidean and second countable,
      but not Hausdorff.
    \item A space $X$ which is Hausdorff and second countable, but not locally
      Euclidean.
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}

% -----------------------------------------------------
% Sixth
% -----------------------------------------------------
\pagebreak

\begin{problem}{6}
  Let $h$ be a continuous real-valued function on
  $S^1 = \{x^2 + y^2 = 1 \subset \mathbb{R}^2\}$ satisfying
  $h(0, 1) = h(1, 0) = 0$ and $h(-x_1, -x_2) = -h(x_1, x_2)$.
  Define a function $f: \mathbb{R}^2 \rightarrow \mathbb{R}$ by \[
    f(x) = \begin{cases}
      ||x||h(x/||x||) & x \not= 0 \\
      0 & x = 0
    \end{cases}
  \]
  \begin{enumerate}[a)]
    \item Show that $f$ is continuous at $(0, 0)$,
      that the partial derivatives of $f$ at $(0, 0)$ are defined,
      and that more generally all directional derivatives of $f$ are defined.

    \item Show that $f$ is not differentiable at $(0, 0)$
      except if $h$ is identically zero.
  \end{enumerate}
\end{problem}

\begin{proof}
\end{proof}

\end{document}
