\documentclass{article}

\input{../../header}

\begin{document}

\title{Math 574: Homework 3}
\author{Peter Kagey}
\date{Friday, November 1, 2019}

\maketitle

% -----------------------------------------------------
% First problem
% -----------------------------------------------------
\begin{problem}{1} Let $A, B \in M_n(K)$ and assume that $AB = BA$.
  \begin{enumerate}[(a)]
    \item Show that $A$ and $B$ are simultaneously upper diagonalizable.
    \item Show that if $A$ and $B$ are diagonalizable, they are simultaneously
      diagonalizable.
  \end{enumerate}
\end{problem}

\begin{proof} ~ % 2.4.8
  \begin{enumerate}[(a)]
    \item If $v$ is an eigenvector of $B$, then $Av$ is also an eigenvector of $B$
    since \[
      A(Bv) = A(\lambda v) = \lambda(Av) = B(Av),
    \] so the eigenspace $E_\lambda$ of $B$ is invariant under multiplication
    by $A$. Thus by conjugating $U^{-1}AU$ to Jordan canonical form, $U^{-1}BU$
    must be upper triangular.
    \item In the case of diagonalizable matrices, part (a) tells us that any
    eigenbasis of $A$ is an eigenbasis of $B$, so
    by writing $U^{-1}AU = D_A$ where the columns of $U$ are this common
    eigenbasis, we also get that $U^{-1}BU = D_B$ is diagonal.

  \end{enumerate}
\end{proof}
% -----------------------------------------------------
% Second problem
% -----------------------------------------------------
\begin{problem}{2} Let $A \in M_n(\R)$
\end{problem}

\begin{proof} ~
  \begin{enumerate}[(a)]
  \item
  % It is sufficient to show that $v^\top AA^\top v \geq 0$. Of course,
  % $(v^\top A)^\top = A^\top v$.
  Let $v$ be a vector, and let $
    A^\top v = \begin{bmatrix} a_1 & a_2 & \dots & a_n \end{bmatrix}.
  $ Notice that $(A^\top v)^\top = v^\top A$, so \begin{align*}
    v^\top AA^\top v
    &= \begin{bmatrix}
      a_1 & a_2 & \dots & a_n
    \end{bmatrix}^\top
    \begin{bmatrix}
      a_1 & a_2 & \dots & a_n
    \end{bmatrix} \\
    &= a_1^2 + a_2^2 + \dots + a_n^2 \\
    &\geq 0.
  \end{align*} Since this argument holds for an arbitrary $v$, $AA^\top$ is
  semipositive definite. Moreover $AA^\top$ equals its own transpose, \[
    (AA^\top)^\top = (A^\top)^\top A^\top = AA^\top,
  \] so $AA^\top$ is symmetric.
  Notice that by singular value decomposition, $A = U\Sigma V^\top$, with $U$
  and $V$ orthogonal and $\Sigma$ diagonal, so \begin{align*}
    AA^\top
      &= U\Sigma V^\top(U\Sigma V^\top)^\top
      = U\Sigma \underbrace{V^\top V}_I\underbrace{\Sigma^\top}_\Sigma U^\top
      = U\Sigma^2 U^\top \\
    A^\top A
      &= (U\Sigma V^\top)^\top U\Sigma V^\top
      = V \underbrace{\Sigma^\top}_\Sigma \underbrace{U^\top U}_I\Sigma V^\top
      = V\Sigma^2V^\top
  \end{align*}
  Therefore $AA^\top$ is similar to $A^\top A$ with \[
    AA^\top
      = U\Sigma^2U^{-1}
      = U\underbrace{V^{-1}A^\top AV}_{\Sigma^2}U^{-1}
      = (UV^{-1})A^\top A (VU^{-1}).
  \]
  \item A counter example is $
    A = \begin{bmatrix} 0 & i \\ 0 & 1\end{bmatrix}
  $. Notice \[
    AA^\top =
    \begin{bmatrix} 0 & i \\ 0 & 1\end{bmatrix}
    \begin{bmatrix} 0 & 0 \\ i & 1\end{bmatrix} =
    \begin{bmatrix} -1 & i \\ i & 1\end{bmatrix}
  \] but \[
    A^\top A =
    \begin{bmatrix} 0 & 0 \\ i & 1\end{bmatrix}
    \begin{bmatrix} 0 & i \\ 0 & 1\end{bmatrix} =
    \begin{bmatrix} 0 & 0 \\ 0 & 0\end{bmatrix}
  \] which have rank $1$ and $0$ respectively, so cannot be similar.
\end{enumerate}
\end{proof}
% -----------------------------------------------------
% Third problem
% -----------------------------------------------------
\begin{problem}{3} ~
  \begin{enumerate}[(a)]
    \item Show that if $A$ is a positive definite Hermitian matrix, then there
    is a unique $B$ that is positive definite Hermitian with $B^2 = A$.
    \item Given an example of a complex square matrix $A$ so that $A \neq B^2$
    for any $B$.
  \end{enumerate}
\end{problem}

\begin{proof} ~
  \begin{enumerate}[(a)]
    \item If $A$ is positive definite Hermitian, then all eigenvalues are real
    and positive, $\lambda_1, \lambda_2, \hdots, \lambda_n$, and $A$ is diagonalizable with $A = UD_AU^{-1}$ by letting $B = U\operatorname{diag}(\sqrt{\lambda_1}, \sqrt{\lambda_2}, \hdots, \sqrt{\lambda_n})U^{-1}$ we see that
    $(UBU^{-1})^2 = UB^2U^{-1} = UD_AU^{-1} = A$.
    \item Let $\displaystyle A = \begin{bmatrix} 0 & 1 \\ 0 & 0\end{bmatrix}$,
    which has rank $1$. For sake of contradiction, assume there exists $B$ with
    $B^2 = A$. Notice that $B$ cannot have rank $2$ (otherwise
    $\det(B)^2 \neq 0$), and $B$ cannot have rank $0$ (because squaring the zero
    matrix results in the zero matrix.)
    Therefore $B$ must have rank $1$ and so its second row must be a linear
    combination of the first. \[
      B^2
        = \begin{bmatrix} a & b \\ ca & cb\end{bmatrix}^2
        = \begin{bmatrix} a^2 + cab & ab + cb^2 \\ ca^2+c^2ab & cab + c^2b^2\end{bmatrix}^2
        = \begin{bmatrix} a(a + cb) & b(a + cb) \\ ac(a + cb) & cb(a + cb)\end{bmatrix}^2
        = \begin{bmatrix} 0 & 1 \\ 0 & 0\end{bmatrix}
        = A
    \] since $b(a + cb) = 1$ and $cb(a + cb) = 0$, this implies that $c = 0$.
    By looking at the first entry, this means that $a^2 = 0 = a$, but this is a
    contradiction because if $a = c = 0$, then $b(a + cb) = 0 \neq 1$.
  \end{enumerate}
\end{proof}

% -----------------------------------------------------
% Fourth problem
% -----------------------------------------------------
\begin{problem}{4} Show that the set of cyclic elements in $M_n(\C)$ form an open dense subset of $M_n(\C)$.
\end{problem}

\begin{proof}
  A matrix is cyclic if its Jordan normal form consists of blocks with distinct
  eigenvalues.
  \\
  There are two things to show: (a) Given any cyclic $A \in M_n(\C)$, there
  exists an open ball around $A$ such that all matrices in the open ball are
  cyclic, and (b) Given a non-cyclic $B \in M_n(\C)$, every open ball around
  $B$ contains a cyclic matrix.
  \begin{enumerate}[(a)]
    \item Given a matrix $A$ where each Jordan block has a distinct eigenvalue,
    changing each entry by less than $\varepsilon$ means that each eigenvalue
    can change by at most $n\varepsilon$. Thus we can choose $\varepsilon$ small
    enough that the eigenvalues remain distinct.
    \item Suppose that $B$ is not cyclic. Then an arbitrarily small change to the
    entries of $B$ will change the eigenvalue corresponding to a Jordan block,
    and therefore $B + \varepsilon a_{ij}$ will be cyclic. (Where $a_{ij}$
    is notation for a matrix that is all zeroes except for a $1$ in the $ij$ entry.)
  \end{enumerate}
\end{proof}
% -----------------------------------------------------
% Fifth problem
% -----------------------------------------------------
\begin{problem}{5} Let $A, B$ be Hermitian matrices with $A$ positive definite.
  \begin{enumerate}[(a)]
    \item Show that if $B$ is positive semidefinite then $A + B$ is positive
    definite.
    \item Show that the largest eigenvalue $\alpha$ of $A$ is
    $\max\set{\langle Av, v \rangle : ||v|| = 1}$.
    \item Show that there exists an invertible matrix $S$ such that
    $SAS^* = I_n$ and $SBS*$ is diagonal.
  \end{enumerate}
\end{problem}

\begin{proof} ~
  \begin{enumerate}[(a)]
    \item Let $v$ be an arbitrary nonzero vector of suitable size. Then since
    $A$ is positive definite and $B$ is positive semidefinite, \[
      v^\top(A + B)v
        = \underbrace{v^\top A v}_{> 0} + \underbrace{v^\top B v}_{\geq 0}
        > 0,
    \] so $A + B$ is positive definite.
    \item Let $\alpha$ be the largest eigenvalue with corresponding unite
    eigenvector $v_\alpha$, and notice that \[
      \max\set{\langle Av, v \rangle : ||v|| = 1} \geq \alpha
    \] since $
      \langle A v_\alpha, v_\alpha\rangle
      = \langle \alpha v_\alpha, v_\alpha\rangle
      = \alpha \underbrace{\langle v_\alpha, v_\alpha\rangle}_1
      = \alpha.
    $
    Because $A$ is Hermitian, we can write an orthonormal basis of eigenvectors, $\set{v_1, v_2, \dots, v_n}$. With corresponding eigenvalues $\lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_n$.
    Then if we write $v = a_1v_1 + a_2v_2 + \dots + a_nv_n$, and $Av = \lambda_1a_1v_1 + \lambda_2a_2v_2 + \dots + \lambda_na_nv_n$, then \begin{align*}
      \ang{v, Av}
        &= \lambda_1a_1^2 + \lambda_2a_2^2 + \hdots + \lambda_na_n^2 \\
        &\leq \lambda_1a_1^2 + \lambda_1a_2^2 + \hdots + \lambda_1a_n^2 \\
        &= \lambda_1(a_1^2 + a_2^2 + \hdots + a_n^2) \\
        &= \lambda_1
    \end{align*}
    with equality when $a_1 = 1$.
    \item
    % Since $A$ and $B$ are Hermitian, they are diagonalizable. Moreover,
    % since $AB = BA$, they are simultaneously diagonalizable: $UAU^{-1}$ and $UBU^{-1}$ are diagonal.
    Since $A$ is positive definite Hermitian, it can be factored as $U \Lambda U^*$
    with $\Lambda = \operatorname{diag}(\lambda_1, \lambda_2, \dots, \lambda_n)$
    and $\lambda_i > 0$.
    Then let $D = D^* = \operatorname{diag}(\lambda_1^{-1/2}, \lambda_2^{-1/2}, \dots, \lambda_n^{-1/2})$ so that $D\Lambda D = I$. Next, letting $S = UD$ \[
      S\Lambda S^* = UD\Lambda D^*U^* = UIU^* = I.
    \]
    The rest follows by Theorem 4.5.17 in Horn.
  \end{enumerate}
\end{proof}
\end{document}
