\addpoints
\setlength{\parindent}{0cm}

\begin{center}

{\Large\textbf{MATH 040 - Final Exam Preview}}\\
\medskip
{\Large{Spring 2020}}

\vskip1cm
\makebox[0.7\textwidth]{\large Last Name:\enspace\hrulefill}
\vskip0.4cm
\makebox[0.7\textwidth]{\large First Name:\enspace\hrulefill}
\vskip0.4cm

\makebox[0.7\textwidth]{\large Signature:\enspace\hrulefill}
\vskip0.4cm

\end{center}
\bigskip
\textbf{Instructions:}
\begin{itemize}
  \item Try to complete the entire exam, but know that you're welcome to stop after 2 hours--even if the exam is not completed. Please record how much time
  you spent on this front page.
  \item Some subset of the questions on this Final Exam Preview will be selected (and perhaps slightly modified) together with some questions from the quizzes and Midterm Preview to make up your Final Exam.
  \item This Final Exam Preview is longer than the actual final exam, so
  don't worry if you don't complete the whole thing in under an hour.
  \item This Preview will be graded as participation only! I will let everyone
  know which problems they got right or wrong, but everyone who completes the
  assignment will receive full credit.
  \item Feel free to use calculators, books, the internet, your instructors, and
  any TAs or tutors. However, books, the internet, and instructors will not be
  available when you take the actual Final Exam.
  \item If anything is unclear, please give feedback by writing your remark next to the question.
  \item Please show as much work as possible, because that allows me to give you the best feedback.
\end{itemize}
\vskip0.4cm
\makebox[0.7\textwidth]{\large Time Spent:\enspace\hrulefill}
\vskip0.4cm
\begin{center}
  \gradetable[h][questions]
\end{center}

\newpage
\vskip1cm
