\addpoints
\setlength{\parindent}{0cm}

\begin{center}

{\Large\textbf{MATH 040 - Midterm Exam Preview}}\\
\medskip
{\Large{Spring 2020}}

\vskip1cm
\makebox[0.7\textwidth]{\large Last Name:\enspace\hrulefill}
\vskip0.4cm
\makebox[0.7\textwidth]{\large First Name:\enspace\hrulefill}
\vskip0.4cm

\makebox[0.7\textwidth]{\large Signature:\enspace\hrulefill}
\vskip0.4cm

\end{center}
\bigskip
\textbf{Instructions:}
\begin{itemize}
  \item Some subset of the questions on this Midterm Exam Preview will be selected (and perhaps slightly modified) to make up your Midterm Exam.
  \item This Preview will be graded as participation only! I will let everyone
  know which problems they got right or wrong, but everyone who completes the
  assignment will receive full credit.
  \item This Midterm Exam Preview is longer than the actual midterm, so
  don't worry if you don't complete the whole thing.
  \item Please spend at least 1 hour attempting the Preview, and you're welcome to stop after 2 hours--even if the exam is not completed. Please record how much time
  you spent on this front page.
  \item Feel free to use calculators, books, the internet, your instructors, and
  any TAs or tutors. However, books, the internet, and instructors will not be
  available when you take the actual Midterm Exam.
  \item You are welcome to give feedback on any questions by leaving a note next to the question.
\end{itemize}
\vskip0.4cm
\makebox[0.7\textwidth]{\large Time Spent:\enspace\hrulefill}(between 1 and 2 hours)
\vskip0.4cm
\begin{center}
  \gradetable[h][questions]
\end{center}

\newpage
\vskip1cm
